```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```



```{css, echo=FALSE}
# set max height of code chunks using the following css styling

pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 500px;
}
```

```{r eval=FALSE}
library(dplyr)
```



### Combine spatial data {-}

Combine spatial data into a single data frame

#### Load data {-}

```{r eval=FALSE}
# bird detection data
load("0_data/manual/bird/bird_callingLake_wide_5.rData")

# QPAD offsets
load("0_data/manual/bird/callingLake_QPAD_offsets.rData")

# LIDAR
load("0_data/manual/spatialCov/LiDAR/ss_lidar_df.rData")

# Classified landcover
load("0_data/manual/spatialCov/CASFRI/pkey_cas_cov.rData")
load("0_data/manual/spatialCov/QPAD/callingLake_QPAD_LC.rData")
```


#### Select covariates {-}

- *CAS-FRI derived*
  - rast_forest_age
  - ss_forest_origDate_mean
  - ss_forest_origDate_sd
  - Forest_Type  

- *From 2005 MODIS Landcover*
  - LCC2
  - LCC4
  
- *LiDAR*  
  - canopy_relief_ratio                      
  - elev_0pnt15_to_2pnt00_return_proportion  
  - elev_10pnt00_to_15pnt00_return_proportion
  - elev_15pnt00_to_20pnt00_return_proportion
  - elev_20pnt00_to_25pnt00_return_proportion
  - elev_25pnt00_to_30pnt00_return_proportion
  - elev_2pnt00_to_4pnt00_return_proportion  
  - elev_30pnt00_to_50pnt00_return_proportion
  - elev_4pnt00_to_6pnt00_return_proportion  
  - elev_6pnt00_to_8pnt00_return_proportion  
  - elev_8pnt00_to_10pnt00_return_proportion 
  - elev_below_0pnt15_return_proportion      
  - elev_cv                                  
  - elev_kurtosis                            
  - elev_maximum                             
  - elev_mean                                
  - elev_p50                                 
  - elev_p95                                 
  - elev_p99                                 
  - elev_stddev                              
  - percentage_first_returns_above_2pnt00    
  - percentage_first_returns_above_mean      
  - total_all_returns       

```{r eval=FALSE}
# LiDAR
l<-ss_lidar_df %>%
  select(SS, canopy_relief_ratio, elev_0pnt15_to_2pnt00_return_proportion, elev_10pnt00_to_15pnt00_return_proportion, elev_15pnt00_to_20pnt00_return_proportion, elev_20pnt00_to_25pnt00_return_proportion, elev_25pnt00_to_30pnt00_return_proportion, elev_2pnt00_to_4pnt00_return_proportion, elev_30pnt00_to_50pnt00_return_proportion, elev_4pnt00_to_6pnt00_return_proportion, elev_6pnt00_to_8pnt00_return_proportion, elev_8pnt00_to_10pnt00_return_proportion , elev_below_0pnt15_return_proportion, elev_cv, elev_kurtosis, elev_maximum , elev_mean, elev_p05 , elev_p10 , elev_p20 , elev_p25 , elev_p30 , elev_p50 , elev_p60 , elev_p70 , elev_p75 , elev_p80 , elev_p95 , elev_p99 , elev_stddev, percentage_first_returns_above_2pnt00, percentage_first_returns_above_mean, total_all_returns)

# CAS-FRI
c<-pkey_cas_cov %>%
  select(SS, PKEY, rast_forest_age, ss_forest_origDate_mean, ss_forest_origDate_sd, Forest_Type)

# MODIS
m<-callingLake_QPAD_LC %>%
  select(SS, PKEY, LCC2, LCC4)

```


#### Join spatial data {-}

```{r eval=FALSE}
spatialCov_all<-left_join(c, m)%>%
  left_join(l)

save(spatialCov_all, file = "2_pipeline/store/spatialCov_all.rData")
```



### Combine observation and spatial data {-}


#### Join data {-}

```{r eval=FALSE}

bd<-distinct(bird_callingLake_wide_5)

data_main<-bd%>%
  dplyr::select(-lidar_date)%>%
  rename(n_survey_years=n_years) %>%
  left_join(callingLake_QPAD_offsets) %>% 
  left_join(spatialCov_all, by="PKEY") %>%
  dplyr::select(-c(SS.y)) %>%
  rename(SS=SS.x) %>%
  distinct()

save(data_main, file = "0_data/manual/data_main.rData")
```

```{r echo=FALSE, message = FALSE, results="asis"}
load("0_data/manual/data_main.rData")

knitr::kable(head(data_main)) %>%
  kable_styling(font_size = 10, bootstrap_options= c("striped", "hover", "condensed" ), full_width= F, position="center") %>%
  column_spec(1, width= "20em")%>%
  scroll_box(width = "100%", height = "400px")
```


### Filter data {-}

Filter by:
- To minimize the influence of forest edges and adjacent differently aged forest, remove stations that are surrounded by a variety of forest ages (sd>5).


```{r eval=FALSE}
load("0_data/manual/data_main.rData")
f1<-data_main%>%
  filter(ss_forest_origDate_sd<11)

nrow(f1 %>%
  dplyr::select("SS.x")%>%
  distinct())


f2<-data_main%>%
  filter(ss_forest_origDate_sd<6)

nrow(f2 %>%
  dplyr::select("SS.x")%>%
  distinct())

f3a<-data_main %>% 
  filter(rast_forest_age > 30 & ss_forest_origDate_sd<11)
f3b<-data_main %>% 
  filter(rast_forest_age < 31 & ss_forest_origDate_sd<6)  
f3c<-rbind(f3a, f3b)


data_main_filter<-f2

save(data_main_filter, file = "2_pipeline/store/data_main_filter.rData")

```



